---
title: "Assignment 1"
author: "Alex Sacharow & Torben Klausa"
date: "October 07, 2016"
output: 
  pdf_document:
    toc: true
    number_sections: true
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
## Setting Working directory
try(setwd("/home/torben/GIT/1st-Pair-Assignment"), silent = TRUE)
try(setwd("D:/Eigene Datein/Dokumente/Uni/Hertie/Materials/Collaborative Social Science Data Analysis/1st-Pair-Assignment"), silent = TRUE)
```


# Introduction

In this assignment we draw two data sets from the GitHub repository [fivethirtyeight](https://github.com/fivethirtyeight) and apply some descriptive statistics on them.[^footnote1] If you have set this file up correctly your working directory should appear here:


```{r show working directory, echo=FALSE}
getwd()
```

# First data set: Police Killings

The first data set we choose is about police killings in the US which was gathered by the Guardian. Information on the structure of the raw data set can be accessed on [GitHub: fivethirtyeight](https://github.com/fivethirtyeight/data/tree/master/police-killings). FiveThirtyEight wrote themselves about it on their [website](http://fivethirtyeight.com/features/where-police-have-killed-americans-in-2015/) and the corresponding Guardian story is available on [The counted](https://www.theguardian.com/us-news/ng-interactive/2015/jun/01/the-counted-police-killings-us-database). What fallows are our own descriptive statistics of the data set.

```{r data, cache=FALSE, include=FALSE}
source("getData.R")
#source("stats+Graphics.R")
```

## Basic one variable descriptive statistics

To begin we will just plot a few basic descriptives using R internal functions. 


# Second data set


[^footnote1]: This is a test footnote. See @Acemoglu2000 [33-40].

# References